A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-14 tsibble.Rmd
done with revision about data pipelines openning

This tiny piece of information will make implicit missing time handlers handy and be in harmony with statistics and models that use seasonal periods.

@@

A data pipeline describes the flow of data through an analysis, and can generally assist in conceptualizing the process, when it is applied to a variety of problems. @unix coined the term "pipelines" in software development while developing Unix at Bell Labs. In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each program becomes the input to another. The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can also be considered a data pipeline. @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot. @xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **XGobi**, **GGobi**, **Orca**, **plumbr** and **cranvas**, respectively. The pipeline is typically described with a one way flow, from data to plot. For interactive graphics, where all plots need to be updated when a user interacts with one plot, the events typically trigger the data pipeline to be run. @xie2014reactive uses a reactive programming framework, to implement the pipeline, in which user's interactions trigger a sequence of modules to update their views, that is, practically the same as running the data pipeline producing each plot.

Building a data pipeline is technically difficult: many implementation decisions have to be made about the interface, input and output objects and functionality. The tidy data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling. Each module communicates with the others, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.

What is notable about an effective implementation of a data pipeline is that it coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies. A fluent pipeline glues tidy data and the grammar of data manipulation together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline. New data tools developed in the R package **tsibble** [@R-tsibble] articulate the time series data pipeline, which shepherds raw temporal data through to time series analysis, and plots. More detailed explanations are given in the following sections, and the examples.

@@

Most importantly, a new ecosystem for tidy time series analysis has been undertaken, using the tsibble framework, and is called "tidyverts", a play on tidyverse that acknowledges the time series analysis purpose.

>>>>

This tiny piece of information will be carried over for tsibble-centric functions, for example to make implicit missing time handling handy and be in harmony with statistics and models that use seasonal periods.

@@

A data pipeline describes the flow of data through an analysis, and can generally assist in conceptualizing the process for a stream of problems. @unix coined the term "pipelines" in software development while developing Unix at Bell Labs. In Unix-based computer operating systems, a pipeline chains together a series of operations based on their standard streams, so that the output of each program becomes the input to another. The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can also be considered a data pipeline. @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot. @xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **XGobi**, **GGobi**, **Orca**, **plumbr** and **cranvas**, respectively.

<!--
The pipeline is typically described with a one way flow, from data to plot. For interactive graphics, where all plots need to be updated when a user interacts with one plot, the events typically trigger the data pipeline to be run. @xie2014reactive uses a reactive programming framework, to implement the pipeline, in which user's interactions trigger a sequence of modules to update their views, that is, practically the same as running the data pipeline producing each plot.
-->

A fluent data pipeline anticipates a standard data structure. The tsibble data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling in temporal contexts. It provides a data infrastructure to a new ecosystem, **tidyverts**, a play on tidyverse that acknowledges the time series analysis purpose, in the spirit of "tidy data" to the **tidyverse**.

@@

It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline. New data tools developed in the R package **tsibble** [@R-tsibble] articulate the temporal data pipeline, which shepherds wild temporal data through to robust analysis, and plots.
